Synthesising Complex Termination Arguments

### 2014-06-03 papers/termination/paper.tex
Restructuring

We can specify the existence of a ranking function, and therefore the termination
of a program, using a second order formula:

@@

\begin{itemize}
 \item There are a TON of programs (way more than you'd expect).
 \item There are a TON of terminating programs, and for our model of computation we can count
  how many (the Chaitin constant).
 \item There are a TON of ranking functions (way more than you'd expect, but not many as a
  fraction of programs).
 \item There are not many linear functions.
 \item Most terminating programs don't admit linear ranking functions.
 \item Kolmogorov complexity is relevant.
\end{itemize}

@@

@@

\section{Solving the Synthesis Formula}

\subsection{Program Synthesis as Second-order Satisfaction}
The program synthesis problem can be described as finding a satisfying assigment to the
synthesis formula:

 $$\exists P: X \to Y . \forall x : X . \sigma(x, P(x))$$
 \end{definition}

\begin{definition}[Generalised Synthesis Formula]
 $$\exists P_1: X_1 \to Y_1, \ldots, P_N: X_N \to Y_N , \vec{x}: X^e . \forall \vec{y}: X^a: X . \sigma(x, P_1, \ldots, P_N, y) $$
\end{definition}

\input{kalashnikov}

\section{Miscellaneous Theory}

\subsection{Termination theory}

@@

\subsection{Synthesis}

@@

\begin{conjecture}
 Most incorrect programs are incorrect on most inputs.
\end{conjecture}

>>>>

@@

\begin{itemize}
 \item There are a TON of programs (way more than you'd expect).
 \item There are a TON of terminating programs, and for our model of computation we can count
  how many (the Chaitin constant).
 \item There are a TON of ranking functions (way more than you'd expect, but not many as a
  fraction of programs).
 \item There are not many linear functions.
 \item Most terminating programs don't admit linear ranking functions.
 \item The Curry-Howard isomorphism
 \item Kolmogorov complexity is relevant for understanding termination proofs.
\end{itemize}

@@

\section{Termination as Second-Order Satisfaction}
We can specify the existence of a ranking function, and therefore the termination
of a program, using a second order formula:

\section{Second-Order Satisfaction as Program Synthesis}
The program synthesis problem can be described as finding a satisfying assigment to the
synthesis formula:

 $$\exists P \cdotp \forall x \cdotp \sigma(x, P(x))$$
 \end{definition}

\begin{definition}[Generalised Synthesis Formula]
 $$\exists P, \vec{x} \cdotp \forall \vec{y} \cdotp \sigma(x, P, y) $$
\end{definition}

@@

@@

\section{Solving the Synthesis Formula}
\input{kalashnikov}

\section{Genetic Programming and Incremental Evolution}

We begin by observing that the asymptotic complexity of all of our synthesis
backends are equal, assuming $P \neq NP$.  This complexity is:

$$O(2^{K(f)}$$

Where $K(f)$ is the Kolmogorov complexity of $f$, which is $O(\log Y^X) = O(X)$
so the complexity is doubly exponential in the width of $X$.

@@
